Inverse energy transfer in three-dimensional quantum vortex flows

### 2025-11-19 main_REV.tex
Final update to manuscript - PZS
Final update of paper
Cleaned the directory

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\NOTE}[1]{\textbf{\color{red}[#1]}}
\newcommand*{\CHANGE}[1]{{\color{magenta}#1}}
\newcommand*{\GK}[1]{{\color{blue}#1}}
\newcommand*{\DEL}[1]{{\color{green}#1}}

@@

To explain the inverse energy transfer shown in Fig.~\ref{fig:spectra} (b),
we look whether the reconnection triggers a chirality imbalance in the flow.  
We decompose the incompressible Fourier modes of the normal fluid velocity 
into helical modes \cite{waleffe-1992}:
\begin{equation}
\hat{\mathbf{v}}_n (\k) = \hat{\mathbf{v}}_n^+(\k) +\hat{\mathbf{v}}_n^-(\k)=
 v_n^+(\mathbf{k}) \mathbf{h}^+(\mathbf{k})+v_n^-(\mathbf{k}) \mathbf{h}^-(\mathbf{k}),
\end{equation}
%
where $\mathbf{h}^\pm (\mathbf{k})$ are the two eigenvectors of the curl 
operator, \textit{i.e.} $i\k~\times~\h^{\pm}(\k)~=~\pm k \h^{\pm}(\k)$. 
Similarly, we decompose the transverse modes of the mutual friction force:
$\hat{\mathbf{F}}_{ns}^{\perp}(\k) = f^+(\k) \mathbf{h}^+ + f^-(\k) \mathbf{h}^-$
(the  Fourier modes of $\mathbf{F}_{ns}$ parallel to the wavemumber 
$\k$ do not play any role in the time evolution of $\mathbf{v}_n$ 
due to the incompressible constraint). Finally, the helical decomposition naturally allow us decompose the total helicity as $\mathcal{H}=\mathcal{H}^+-\mathcal{H}^-$ \cite{plunianInverseCascadeEnergy2020a}.
% %
% The spectral energy densities corresponding to the helical modes are
% $E^{\pm}(\k) = (1/2) |v_n^\pm(\k)|^2$, the total spectral density 
% being $E(\k) = E^{+}(\k) + E^{-}(\k)$. Similarly the spectral helicity 
% density is 
% \begin{equation}
% \begin{split}
% H(\k) & = (1/2) \hat{\mathbf{v}}_n (\k)\!\!\! ~\cdot~\!\!\! \hat{\bm{\omega}}_n^* (\k) = \\
% & = k E^+(\k) - k E^-(\k)  = H^+(\k) - H^-(\k),
% \end{split}
% \end{equation}
% where $\bm{\omega}_n$ is the normal fluid vorticity, the star indicates 
% complex conjugate and $H^\pm$ is the helicity contribution of each 
% separate helical mode. \NOTE{GK: the standard definition of the energy and helicity spectra contains the sum over the angles, so the previous definition are wrong (except if you redifined them for some reason).}
%The rate of change $I^\pm(\k)$ of the energies $E^{\pm}(\k)$ arising from the mutual friction is proportional to the corresponding helical coefficients $f^\pm$, \textit{i.e.} $I^\pm(\k) = \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$, and the related signed helicity injection is $I_H^\pm(\k) = k \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$.

A chiral imbalance occurs if the mutual friction force is helical, 
\textit{i.e.} if the ratio $|f^+|^2/|f^-|^2 \neq 1$, with $|f^\pm|^2$ the total squared norm of the mutual friction components. 
In Fig.~\ref{fig:mutual-friction-decomp}, we show the temporal evolution of $|f^+|^2/|f^-|^2$ for both temperatures, for orthogonal reconnections
(Hopf-links reconnections show a similar behaviour \cite{suppMat}). 
%
\begin{figure}[h!]
    \centering
    \includegraphics*[width=0.48\textwidth]{fmfDecompFig.pdf}
    \caption{Temporal evolution of the ratio of helically-projected mutual friction force components $f^\pm$, for orthogonal reconnections. 
    Inset: temporal evolution of total helical components.}.
    \label{fig:mutual-friction-decomp}
\end{figure}
%
It is apparent
that during and after the reconnection, the mutual friction force is strongly chiral,
injecting more negative helicity than positive helicity. 
As a result, the ratio $\mathcal{H}^+/\mathcal{H}^-$ 
(reported in the inset of Fig.~\ref{fig:mutual-friction-decomp}) 
decreases significantly at reconnection and remains smaller than unity even at later times, 
indicating that the flow is persistently chiral. We conclude that the reconnection triggers indeed a chiral imbalance, \CHANGE{leading to the observed inverse energy transfer}.
For the sake of completeness, it is worth noting that the total helicity of the flow (obtained by adding the superfluid helicity to that of the normal fluid $\mathcal{H}$) is not conserved.
From Fig.~\ref{fig:mutual-friction-decomp}, we consider a time $t^*$ after reconnection where the ratio $|f^+|^2/|f^-|^2$ has sufficiently decayed to a roughly 5\% difference and we determine the non-dimensional timescale $\tau = (t^*-t_R)/\tau_R$ during which the mutual friction force is chiral as a result of reconnections: $\tau \approx 0.01$ and $\tau \approx 0.005$
for $T=1.9$K and $T=2.1$K, respectively, corresponding dimensionally to $\tau \approx 0.1$s for both temperatures. 
In superfluid turbulence, the timescale between two consecutive reconnections can be smaller than $\tau$ provided that the
vortex line density $\mathcal{L}$ (length of vortices per unit volume) is larger than $10^8 \text{m}^{-2}$ \cite{stasiak2025experimental,barenghi2004},
a condition which is easily met in superfluid helium experiments \cite{roche2007,Babuin2014}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\NOTE}[1]{\textbf{\color{black}[#1]}} %Red
\newcommand*{\CHANGE}[1]{{\color{black}#1}}    %Magenta
\newcommand*{\GK}[1]{{\color{black}#1}}           %Blue
\newcommand*{\DEL}[1]{{\color{black}#1}}         %Green

@@

To explain the inverse energy transfer shown in Fig.~\ref{fig:spectra} (b),
we look whether the reconnection triggers a chirality imbalance in the flow.  
We decompose the incompressible Fourier modes of the normal fluid velocity 
into helical modes \cite{waleffe-1992}:
\begin{equation}
\hat{\mathbf{v}}_n (\k) = \hat{\mathbf{v}}_n^+(\k) +\hat{\mathbf{v}}_n^-(\k)=
 v_n^+(\mathbf{k}) \mathbf{h}^+(\mathbf{k})+v_n^-(\mathbf{k}) \mathbf{h}^-(\mathbf{k}),
\end{equation}
%
where $\mathbf{h}^\pm (\mathbf{k})$ are the two eigenvectors of the curl 
operator, \textit{i.e.} $i\k~\times~\h^{\pm}(\k)~=~\pm k \h^{\pm}(\k)$. 
Similarly, we decompose the transverse modes of the mutual friction force:
$\hat{\mathbf{F}}_{ns}^{\perp}(\k) = f^+(\k) \mathbf{h}^+ + f^-(\k) \mathbf{h}^-$
(the  Fourier modes of $\mathbf{F}_{ns}$ parallel to the wavemumber 
$\k$ do not play any role in the time evolution of $\mathbf{v}_n$ 
due to the incompressible constraint). Finally, the helical decomposition naturally allow us decompose the total helicity as $\mathcal{H}=\mathcal{H}^+-\mathcal{H}^-$ \cite{plunianInverseCascadeEnergy2020a}.
% %
% The spectral energy densities corresponding to the helical modes are
% $E^{\pm}(\k) = (1/2) |v_n^\pm(\k)|^2$, the total spectral density 
% being $E(\k) = E^{+}(\k) + E^{-}(\k)$. Similarly the spectral helicity 
% density is 
% \begin{equation}
% \begin{split}
% H(\k) & = (1/2) \hat{\mathbf{v}}_n (\k)\!\!\! ~\cdot~\!\!\! \hat{\bm{\omega}}_n^* (\k) = \\
% & = k E^+(\k) - k E^-(\k)  = H^+(\k) - H^-(\k),
% \end{split}
% \end{equation}
% where $\bm{\omega}_n$ is the normal fluid vorticity, the star indicates 
% complex conjugate and $H^\pm$ is the helicity contribution of each 
% separate helical mode. \NOTE{GK: the standard definition of the energy and helicity spectra contains the sum over the angles, so the previous definition are wrong (except if you redifined them for some reason).}
%The rate of change $I^\pm(\k)$ of the energies $E^{\pm}(\k)$ arising from the mutual friction is proportional to the corresponding helical coefficients $f^\pm$, \textit{i.e.} $I^\pm(\k) = \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$, and the related signed helicity injection is $I_H^\pm(\k) = k \mathrm{Re} [f^\pm (v_n^\pm)^*] /\rho_n$. 
A chiral imbalance occurs if the mutual friction force is helical, 
\textit{i.e.} if $\mathcal{F}_R \neq 1$, where
\begin{equation}
	\mathcal{F}_R(t) = \frac{\int dk |f^+(k,t)|^2}{\int dk |f^-(k,t)|^2},
\end{equation}
with $|f^\pm|^2$ the squared norm of the mutual friction
helical decomposition coefficients. 
In Fig.~\ref{fig:mutual-friction-decomp}, we show the temporal evolution of $\mathcal{F}_R$ for both temperatures, for orthogonal reconnections
(Hopf-links reconnections show a similar behaviour \cite{suppMat}). 
%
\begin{figure}[h!]
    \centering
    \includegraphics*[width=0.48\textwidth]{fmfDecompFig-1.pdf}
    \caption{Temporal evolution of the ratio of helically-projected mutual friction force components $f^\pm$, for orthogonal reconnections. 
    Inset: temporal evolution of total helical components.}
    \label{fig:mutual-friction-decomp}
\end{figure}
%
It is apparent
that during and after the reconnection, the mutual friction force is strongly chiral,
injecting more negative helicity than positive helicity. 
As a result, the ratio $\mathcal{H}^+/\mathcal{H}^-$ 
(reported in the inset of Fig.~\ref{fig:mutual-friction-decomp}) 
decreases significantly at reconnection and remains smaller than unity even at later times, 
indicating that the flow is persistently chiral. We conclude that the reconnection triggers indeed a chiral imbalance, \CHANGE{leading to the observed inverse energy transfer}.
For the sake of completeness, it is worth noting that the total helicity of the flow (obtained by adding the superfluid helicity to that of the normal fluid $\mathcal{H}$) is not conserved.
From Fig.~\ref{fig:mutual-friction-decomp}, we consider a time $t^*$ after reconnection where the ratio $\mathcal{F}_R$ has sufficiently decayed to a roughly 5\% difference and we determine the non-dimensional timescale $\tau = (t^*-t_R)/\tau_R$ during which the mutual friction force is chiral as a result of reconnections: $\tau \approx 0.01$ and $\tau \approx 0.005$
for $T=1.9$K and $T=2.1$K, respectively, corresponding dimensionally to $\tau \approx 0.1$s for both temperatures. 
In superfluid turbulence, the timescale between two consecutive reconnections can be smaller than $\tau$ provided that the
vortex line density $\mathcal{L}$ (length of vortices per unit volume) is larger than $10^8 \text{m}^{-2}$ \cite{stasiak2025experimental,barenghi2004},
a condition which is easily met in superfluid helium experiments \cite{roche2007,Babuin2014}.
